---
title: "03_Model_Agreement"
author: "Matthew Coghill"
date: "7/14/2020"
output: html_document
---

```{r}

ls <- c("tidyverse", "terra", "foreach", "raster")
new.packages <- ls[!(ls %in% installed.packages()[, "Package"])]
if(length(new.packages))
  install.packages(new.packages)
lapply(ls, library, character.only = TRUE)[0]
rm(ls, new.packages)

```



```{r Directory setup}

AOI <- "Sechelt"
AOI_dir <- file.path(".", paste0(AOI, "_AOI"))
map_res <- 4

model_dir <- file.path(AOI_dir, "2_map_predictions")
covariate_folder <- file.path(AOI_dir, "1_map_inputs", "covariates")
res_folder <- paste0(map_res, "m")
dsmart_folder <- file.path(AOI_dir, "1_map_inputs", "dsmart")

covariates <- c(
  list.files(file.path(covariate_folder, res_folder), full.names = TRUE, pattern = ".tif$"),
  list.files(file.path(dsmart_folder, "simple"), pattern = ".tif$", full.names = TRUE),
  list.files(file.path(dsmart_folder, "output", "probabilities"), pattern = ".tif", full.names = TRUE),
  file.path(dsmart_folder, "output", "realisations", "realisation_1.tif")
)

```

Simple method proposed by Mike: Count the number of maps that predicted a plant at a given pixel. This requires a bit of adjustment as well: First, reclassify all map probability values below 0.5 to be 0, and then count the maps that predict anything above that value. Repeat for above and below 0.75.

```{r Pres Abs model agreement}

pred_folder <- file.path(model_dir, "pres_abs_predictions")
out_folder <- file.path(model_dir, "pres_abs_agreement")
dir.create(out_folder, showWarnings = FALSE)
i <- "CORNCAN"
j <- 0.5
for(i in c("CORNCAN", "RUBUSPE")) {
  model_stack <- rast(
    foreach(x = dir(pred_folder, pattern = i, full.names = TRUE), .combine = c) %do% {
      list.files(x, pattern = "_TRUE.tif$", 
                 full.names = TRUE)
    })
  
  for(j in c(0.5, 0.75)) {
    # Classify cell values
    rcl <- classify(model_stack, matrix(c(0, j, 0, j, 1, 1), ncol = 3, byrow = TRUE))
    
    # Create summed raster
    agreement <- app(
      rcl, fun = "sum", 
      filename = file.path(out_folder, paste0(i, "_pres_abs_gt_", 100 * j, "_model_agreement.tif")), 
      overwrite = TRUE, wopt = list(datatype = "INT2S"))
    
    agreement_freq <- as.data.frame(freq(agreement, bylayer = FALSE)) %>%
      magrittr::set_colnames(c("zone", "count"))
    
    # Zonal statistics
    # z_stats <- lapply(covariates, function(x) {
    #   foreach(k = c("mean", "sd", "min", "max", "sum"), .final = function(y) {
    #     Reduce(function(...) merge(..., by = "zone"), y)
    #   }) %do% {
    #     zonal(rast(x), agreement, k, na.rm = TRUE) %>%
    #       magrittr::set_colnames(c("zone", k))
    #   } %>%
    #     merge(agreement_freq, by = "zone")
    # }) %>% magrittr::set_names(basename(tools::file_path_sans_ext(covariates)))
    
    # Terra package top level notation, since zonal function is broken for 
    # large rasters. This works fine though
    z_stats <- foreach(k = covariates, .final = function(y) {
      Reduce(function(...) merge(..., by = "zone"), y)
    }) %do% {
      terra:::.getSpatDF(rast(k)@ptr$zonal(agreement@ptr, "mean", na.rm = TRUE))
    } %>% merge(agreement_freq, by = "zone")
    
    # Cleanup
    tmpFiles(remove = TRUE)
  }
}

```

I thought of trying to do a similar technique with cover. Because of really small decimal places in raw cover values, I figured I should reclassify cover values less than 1% as a 0 value, and then everything else is a value of 1 in order to count where models agree that there is cover greater than 1%

```{r Cover model agreement}

pred_folder <- file.path(model_dir, "cover_predictions")
out_folder <- file.path(model_dir, "cover_agreement")
dir.create(out_folder, showWarnings = FALSE)

for(i in c("CORNCAN", "RUBUSPE")) {
  model_stack <- rast(
    foreach(x = dir(pred_folder, pattern = i, full.names = TRUE), .combine = c) %do% {
      list.files(x, pattern = "_cover.tif$", 
                 full.names = TRUE)
    })
  
  for(j in c(1, 5)) {
    rcl <- classify(model_stack, matrix(c(0, j, 0, j, 100, 1), ncol = 3, byrow = TRUE))
    agreement <- as.factor(sum(rcl))
    writeRaster(agreement, 
                file.path(out_folder, paste0(i, "_cover_gt_", j, "_model_agreement.tif")), 
                overwrite = TRUE, wopt = list(datatype = "INT2S"))
    tmpFiles(remove = TRUE)
  }
}

```

The outputs of the above calculations are "number of maps that agree that plant x has a probability of presence/has a cover percentage greater than y". These numbers of maps can be summed, and it's a weird metric but it is essentially the number of models that agree that a plant would be present and that the cover would be greater than 5%

```{r}

pres_folder <- file.path(model_dir, "pres_abs_agreement")
cover_folder <- file.path(model_dir, "cover_agreement")
cover_vals <- c(1, 5)
pres_probs <- c(50, 75)
for(i in c("CORNCAN", "RUBUSPE")) {
  for(j in cover_vals) {
    for(k in pres_probs) {
      model_stack <- rast(c(list.files(pres_folder, pattern = paste0(i, ".*.", k), full.names = TRUE), 
                            list.files(cover_folder, pattern = paste0(i, ".*.", j), full.names = TRUE))) %>% 
        magrittr::set_names(c("pres_abs", "cover"))
      
      # Only want to keep high cell values, indicating that many models agree at a 
      # given raster pixel
      model_mask <- foreach(l = 1:nlyr(model_stack), .final = rast) %do% {
        n_layers <- max(minmax(model_stack[[l]]))
        classify(model_stack[[l]], matrix(c(0, floor(n_layers / 2), 0, 
                                            floor(n_layers / 2), Inf, 1), ncol = 3, byrow = TRUE))
      } %>% 
        sum() %>% 
        classify(matrix(c(0, 2, 0, 2, Inf, 1), ncol = 3, byrow = TRUE))
      
      # Add both pres_abs and cover rasters, mask out areas defined from the masking
      # function above, and finally only keep cell values that are minimum of the
      # mean of the number of agreeing models...
      model_sum <- sum(model_stack) %>% 
        mask(model_mask, maskvalue = 0, updatevalue = 0) %>% 
        classify(matrix(c(0, floor(mean(minmax(model_stack)["rmax", ])), 0), 
                        ncol = 3, byrow = TRUE),
                 filename = file.path(model_dir, paste0(i, "_cover_gt_", j, "_pres_abs_gt_", k, "_agreement.tif")), 
                 overwrite = TRUE, wopt = list(datatype = "INT2S"))
    }
  }
  
  # Produce a map of likely locations to find a given plant based on all other maps produced
  out_maps <- rast(list.files(model_dir, pattern = paste0(i, ".*..tif$"), full.names = TRUE))
  likely_locations <- classify(out_maps, matrix(c(0, 1, 0, 1, Inf, 1), ncol = 3, byrow = TRUE)) %>% 
    sum() %>% 
    classify(matrix(c(0, floor(nlyr(out_maps) / 2), 0), ncol = 3, byrow = TRUE), 
             filename = file.path(model_dir, paste0(i, "_likely_locations.tif")), 
                 overwrite = TRUE, wopt = list(datatype = "INT2S"))
      
  tmpFiles(remove = TRUE)
  
}

# Attempt to find areas that all plants overlap based on the previous maps.
likely_files <- list.files(model_dir, pattern = "_likely_locations.tif", full.names = TRUE)
efficiency <- rast(likely_files) %>% 
  classify(matrix(c(0, 1, 0, 1, Inf, 1), ncol = 3, byrow = TRUE)) %>% 
  sum() %>% 
  classify(matrix(c(0, length(likely_files), 0, length(likely_files), Inf, 1), ncol = 3, byrow = TRUE), 
           filename = file.path(model_dir, "efficient_locations.tif"), 
                 overwrite = TRUE, wopt = list(datatype = "INT2S"))
      
tmpFiles(remove = TRUE)

```

